\chapter{General Introduction}
\label{chap:intro}

\section{Problem Statement}
In the modern digital landscape, network security has become a paramount concern for organizations of all sizes. The proliferation of connected devices and the increasing sophistication of cyber-attacks have made traditional security measures, such as firewalls, insufficient to protect against all types of threats. Network Intrusion Detection Systems (NIDS) have emerged as a critical layer of defense, providing visibility into network traffic and identifying malicious patterns that might otherwise go unnoticed.

\section{Objectives}
The primary objective of this project is to design and implement a custom NIDS that addresses several key challenges:
\begin{itemize}
    \item \textbf{Real-time Analysis:} Ability to capture and process network packets as they arrive.
    \item \textbf{Modularity:} A system architecture that separates the capture engine from the analysis logic and the user interface.
    \item \textbf{User-Friendly Interface:} A dashboard that provides clear and actionable information to security analysts.
\end{itemize}

\section{Thesis Structure}
This report is organized into four main chapters:
\begin{enumerate}
    \item \textbf{Chapter 1: Background and State of the Art} - Discusses the theoretical foundations of network security and existing IDS solutions.
    \item \textbf{Chapter 2: Related Work} - Reviews previous research and commercial tools in the field of intrusion detection.
    \item \textbf{Chapter 3: Methodology and Design} - Details the architecture and implementation of our proposed system.
    \item \textbf{Chapter 4: Results and Evaluation} - Presents the experimental results and performance analysis.
\end{enumerate}
